Using vector space models based on LDA and LSA for automatically grading exam questions

### 2018-06-18 tex/article.tex
Finished another complete draft of the discussion (deleted the conclusion again and integrated it at the end of the discussion).

An alternative approach is described in \citeasnoun{foltz1998}, which may provide a better solution for future research. Rather than comparing a student's answer to one specific reference answer, the to-be-graded answer is compared to a set of answers already graded by the lecturer. Then, you take the (for example) ten graded answers with the highest similarity score to the to-be-graded answer. The predicted grade will be the weighted average of the grades of the ten most similar answers (weighted by their similarity score).

Another limitation of the current approach (as well as the alternative approach from \citeasnoun{foltz1998}) is that creative answers are punished, such as when a student uses a unique example. This makes the student's answer more dissimilar from the reference answer (or from other students' answers), and results in a lower grade. One solution could be to create a list with terms that are relevant to the question (e.g., [`Whorf', `linguistic', `relativity']) and first reduce the student's answer to only the sentences that contain one of these terms. Sentences such as `The Inuit have many words for snow' would no longer be part of the semantic similarity calculation, perhaps wrongfully lowering the predicted grade.

Even given the limitations described above, a model does not have to be perfect in order to be of potential use. For example, upon inspecting the individual models' results, I could see that whenever the model had assigned a zero to a student answer, the lecturer had also assigned a zero. Thus, the models \textit{are} good enough to automatically detect which answers are worth zero points. A lecturer could use this information to focus his/her attention only on those answers that have not been predicted a grade of zero by the model.

\subsection{Suggestions for future research}
The evaluation metric in this study was the correlation between the assigned and predicted grades. It would be interesting to consider other metrics too, for example the sum of squared errors (SSE) between the assigned and predicted grades. This would enable us to also evaluate a baseline model that would assign the most common grade (in this case 10) to all student answers. Such a baseline model currently could not be employed because a correlation coefficient cannot be calculated when there is no variance in the data.

\section{Conclusion}
While it was good to see that all of the models yielded a statistically significant correlation between predicted and assigned grades, it still seems too early to justify their actual implementation in higher education. We would need to know what is the correlation between two human graders, and equal or go beyond that correlation with our model. Furthermore, the model needs to be able to assign the maximum grade of 10 to excellent answers that are different from the reference answer. With the suggestions proposed in the discussion, we should continue to try to find a reliable method for automatizing the grading process. This could be of great benefit to students, as they could be provided with feedback on their learning process much more quickly.

>>>>

An alternative approach is described in \citeasnoun{foltz1998}, which may provide a better solution for future research. Rather than comparing a student's answer to one specific reference answer, the to-be-graded answer is compared to a set of answers already graded by the lecturer. Then, you take the (for example) ten graded answers with the highest similarity score to the to-be-graded answer. The predicted grade will be the weighted average of the grades of the ten most similar answers (weighted by their similarity score). Another alternative is to use a variety of different reference answers. These could even be obtained automatically by applying automatic paraphrasing to the original reference answer.

The approaches discussed so far, including the one by \citeasnoun{foltz1998}, are all limited in the fact that creative answers are punished, such as when a student uses a unique example. This makes the student's answer more dissimilar from the reference answer (or from other students' answers), and would result in a lower grade. One solution could be to create a list with terms that are relevant to the question (e.g., [`Whorf', `linguistic', `relativity']) and first reduce the student's answer to only the sentences that contain one of these terms. Sentences such as `The Inuit have many words for snow' would no longer be part of the semantic similarity calculation, perhaps wrongfully lowering the predicted grade.

Even given the limitations described above, a model does not have to be perfect in order to be of potential use. For example, upon inspecting the individual models' results, I could see that whenever the model had assigned a zero to a student answer, the lecturer had also assigned a zero. Thus, the current models \textit{are} good enough to automatically detect which answers are worth zero points. A lecturer could use this information to focus his/her attention only on those answers that have not been predicted a zero grade by the model.

Before automatic grading models could actually be implemented in higher education for grading \textit{all} students' answers, we would need to know what is the correlation between two human graders, and equal or go beyond that correlation with our best model. Furthermore, the model needs to be able to assign the maximum grade of 10 to excellent answers that are different from the reference answer. With the suggestions proposed in this discussion, we should continue to try to find a reliable method for automatizing the grading process. This could be of great benefit to students, as they could be provided with feedback on their learning process much more quickly.
